Syntax and Semantics of Abstract Binding Trees

### 2015-12-31 paper.tex
rework sheaf theory preliminaries a bit (#42)

\newcommand\Sheaves[1]{\mathbf{Sh}\left(#1\right)}
\newcommand\Presheaves[1]{\mathbf{Psh}\left(#1\right)}
\newcommand\Sheafify[1]{#1^{\#}}
\newcommand\ConstPsh[1]{\Delta[#1]}
\newcommand\ConstSh[1]{\Delta^\#[#1]}
\newcommand\SCtxSite{\SymSets[\Sorts]}

@@

When
  \begin{tikzcd}[cramped]
    \IMode{\Upsilon}\arrow[r, hook, "\IMode{\varrho}"] & \IMode{\Upsilon'}
  \end{tikzcd}
such that $\Supports{\Upsilon}{\varrho}{m}$ for some $\Member{m}{X(\Upsilon')}$
, we would
expect that we can work backward to a unique $\Of{m'}{\Upsilon}$ such that
$\IsEq{m'\cdot\varrho}{m}$; whilst it is not in general the case for presheaves
$X$, this is precisely the sheaf condition for copresheaves on $\SCtx$
under the atomic topology \cite[p.~126]{mac-lane-moerdijk:1992}.
%%% TODO: add definition of atomic coverage
The sheaf condition, then, ensures that the notion of support is well-behaved in the
following sense: if some term lies in two different fibers, then it may be
\emph{strengthened} uniquely into the fibre of its support.

\paragraph{Sites and coverages}

We will briefly discuss the basics of sheaves on a site. A \emph{site} is a
pair $\Pair{\mathcal{C}}{J}$, with $\mathcal{C}$ a category of ``worlds'' and
$J$ a Grothendieck topology (or ``coverage''). A coverage is a predicate that
picks out for each object $\Of{C}{\mathcal{C}}$, the \emph{covering sieves} of
$C$, subject to several axioms (see \cite{mac-lane-moerdijk:1992} for details).

\paragraph{Covering sieves and matching families}

A \emph{sieve} on an object $C$ is a subfunctor of the Yoneda embedding
$\Define{\Yoneda{C}}{\Hom{\mathcal{C}}{-}{C}}$. We say that such a sieve $S$
covers $C$ just when $\Member{S}{J(C)}$.  For a presheaf $X$ on the site
$\Pair{\mathcal{C}}{J}$, a \emph{matching family} for the sieve $S$ is a
natural transformation $\Of{\phi}{S\ra X}$. An \emph{amalgamation} for $\phi$
is an object $\Member{m}{X(C)}$ such that $\IsEq{m\cdot f}{\phi_D(f)}$ for each
$\Member{f}{S(D)}$, for all objects $D$.

\paragraph{Separation and sheafhood}
Then, we say that the presheaf $X$ is \emph{separated} in case for every sieve
$\Member{S}{J(C)}$ and matching family $\Of{\phi}{S\ra X}$, there is at most
one amalgamation $\Member{m}{X(C)}$ for $\phi$. Additionally, $X$ is called a
\emph{sheaf} when there is \emph{exactly one} such amalgamation.

\paragraph{The atomic coverage}

The atomic coverage $J_\mathsf{atm}$ is the one in which all non-empty sieves
cover. A sieve $\IMode{S}\sqsubseteq\IMode{\Yoneda{C}}$ is non-empty when the
collection $\bigcup_D S(D)$ of all its arrows is non-empty.

In the future, we will write $\SCtxSite$ for the Grothendieck site
$\Pair{\OpCat{\left(\SCtx\right)}}{J_\mathsf{atm}}$, where $J_\mathsf{atm}$ is
the atomic coverage. Sheafhood on the atomic site $\SCtxSite$ amounts to the
following condition, using Definition~\ref{def:support}:
\begin{quote}
  A copresheaf $\Of{X}{\Sets^\SCtx}$ is a sheaf on $\SCtxSite$ just when, for all $\Member{n}{X(\Upsilon')}$, if we have
  \begin{tikzcd}[cramped]
    \IMode{\Upsilon}\arrow[r, hook, "\IMode{\varrho}"] & \IMode{\Upsilon'}
  \end{tikzcd}
  and $\Supports{\Upsilon}{\varrho}{n}$, then there is a unique
  $\Member{m}{X(\Upsilon)}$ such that $\IsEq{m\cdot\varrho}{n}$.
\end{quote}

@@

In addition to the unisorted nominal sets framework, Pitts also briefly
discusses a multi-sorted version of the apparatus in which the countably
infinite set of atoms $\mathbb{A}$ is equipped with a sort assignment
$\Of{\mathfrak{s}}{\mathbb{A}\to\Sorts}$ such that all the fibres of
$\mathfrak{s}$ are countably infinite; then, he defines for any such assignment
the category $\mathbf{Nom}_\mathfrak{s}$ of $\mathfrak{s}$-sorted nominal sets
as a subcategory of the category of presheaves on the group of
$\mathfrak{s}$-respecting permutations on $\mathbb{A}$,
$\Perm_\mathfrak{s}\mathbb{A}$.

>>>>

\newcommand\Sheaves[1]{\mathbf{Sh}\left(#1\right)}
\newcommand\Presheaves[1]{\mathbf{Psh}\left(#1\right)}
\newcommand\Sheafify[1]{#1^{\#}}
\newcommand\ConstPsh[1]{\Delta[#1]}
\newcommand\ConstSh[1]{\Delta^\#[#1]}
\newcommand\SCtxSite{\SymSets[\Sorts]}
\newcommand\AtomicCoverage{J_{\mathtt{atm}}}

@@

When
  \begin{tikzcd}[cramped]
    \IMode{\Upsilon}\arrow[r, hook, "\IMode{\varrho}"] & \IMode{\Upsilon'}
  \end{tikzcd}
such that $\Supports{\Upsilon}{\varrho}{m}$ for some $\Member{m}{X(\Upsilon')}$,
we would expect that we can work backward to a unique $\Of{m'}{\Upsilon}$
such that $\IsEq{m'\cdot\varrho}{m}$; whilst it is not in general the case for
presheaves $X$, this is precisely the sheaf condition for copresheaves on
$\SCtx$ under the atomic topology \cite[p.~126]{mac-lane-moerdijk:1992}, which
we will define in Definition~$\ref{def:atomic-topology}$.

The sheaf condition ensures that the notion of support is well-behaved in the
following sense: if some term lies in two different fibers, then it may be
\emph{strengthened} uniquely into the fiber of its support.

\begin{definition}[Sieves and matching families]
  A \emph{sieve} $S$ on an object $C$ is a subfunctor of the Yoneda embedding
  $\Define{\Yoneda{C}}{\Hom{\mathcal{C}}{-}{C}}$, i.e.\
  $\IMode{S}\sqsubseteq\IMode{\Yoneda{C}}$. For a presheaf $X$ on the category
  $\mathcal{C}$, a \emph{matching family} for a sieve $S$ is a natural
  transformation $\Of{\phi}{S\ra X}$. An \emph{amalgamation} for $\phi$ is an
  object $\Member{m}{X(C)}$ such that $\IsEq{m\cdot f}{\phi_D(f)}$ for each
  $\Member{f}{S(D)}$, for all objects $D$.
\end{definition}

\begin{definition}[Sites and coverages]
  A \emph{site} is a pair $\Pair{\mathcal{C}}{J}$, with $\mathcal{C}$ a category
  of ``worlds'' and $J$ a Grothendieck topology (or ``coverage''). A Grothendieck
  topology $J$ (or \emph{coverage}) is a family of sets of sieves, indexed by
  $\mathcal{C}$-objects; that is, for each object $\Of{C}{\mathcal{C}}$, $J(C)$
  is the set of $C$-sieves which will be said to $\emph{cover}$ $C$. A topology
  $J$ must satisfy several axioms (see \cite{mac-lane-moerdijk:1992} for
  details).
\end{definition}

\begin{definition}[Separation and sheafhood]
  A presheaf on a site $\Pair{\mathcal{C}}{J}$ is a presheaf on $\mathcal{C}$; we
  say that such a presheaf $X$ is \emph{separated} in case for every sieve
  $\Member{S}{J(C)}$ and matching family $\Of{\phi}{S\ra X}$, there is at most
  one amalgamation $\Member{m}{X(C)}$ for $\phi$. Additionally, $X$ is a
  \emph{sheaf} when there is \emph{exactly one} such amalgamation.
\end{definition}

\begin{definition}[The atomic topology]
  \label{def:atomic-topology}
  The atomic coverage $\AtomicCoverage$ is the one in which all non-empty sieves
  cover; in other words:
  \[
    \Define{
      \AtomicCoverage(C)
    }{
      \MkSet{
        S\sqsubseteq\Yoneda{C}
        \mid
        \textstyle
        \bigcup_D S(D)\ \textit{non-empty}
      }
    }
  \]
\end{definition}

In the future, we will write $\SCtxSite$ for the Grothendieck site
$\Pair{\OpCat{\left(\SCtx\right)}}{\AtomicCoverage}$.
Sheafhood on the atomic site $\SCtxSite$ amounts to the
following condition, using Definition~\ref{def:support}:
\begin{quote}
  A copresheaf $\Of{X}{\Sets^\SCtx}$ is a sheaf on $\SCtxSite$ just when, for all $\Member{n}{X(\Upsilon')}$, if we have
  \begin{tikzcd}[cramped]
    \IMode{\Upsilon}\arrow[r, hook, "\IMode{\varrho}"] & \IMode{\Upsilon'}
  \end{tikzcd}
  and $\Supports{\Upsilon}{\varrho}{n}$, then there is a unique
  $\Member{m}{X(\Upsilon)}$ such that $\IsEq{m\cdot\varrho}{n}$.
\end{quote}

@@

In addition to the unisorted nominal sets framework, Pitts also briefly
discusses a multi-sorted version of the apparatus in which the countably
infinite set of atoms $\mathbb{A}$ is equipped with a sort assignment
$\Of{\mathfrak{s}}{\mathbb{A}\to\Sorts}$ such that all the fibers of
$\mathfrak{s}$ are countably infinite; then, he defines for any such assignment
the category $\mathbf{Nom}_\mathfrak{s}$ of $\mathfrak{s}$-sorted nominal sets
as a subcategory of the category of presheaves on the group of
$\mathfrak{s}$-respecting permutations on $\mathbb{A}$,
$\Perm_\mathfrak{s}\mathbb{A}$.
